Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-09-06 inst/doc/nonparametric-bayes.Rmd
Updated nonparametric-bayes.Rmd

This problem is keenly felt in any system in which a manager may fear
crossing a threshold, or tipping point. The tipping point, if it exists 
at all, will lay outside the observed range of the observed data.

@@

Decision making under uncertainty is a ubiquitous challenge of natural 
resource management and conservation.  
Here we illustrate how a stochastic dynamic programming algorithm can 
be driven by the predictions from a Gaussian process model, sidestepping
the need for an accurate model-based description of the system dynamics.

>>>>

This problem is keenly felt in any system where a potential threshold
or tipping point is a concern.  Such a tipping point, if it exists 
at all, will lay outside the observed range of the observed data.

@@

Decision making under uncertainty is a ubiquitous challenge of natural 
resource management and conservation.  
Here we illustrate how a stochastic dynamic programming (SDP) algorithm can 
be driven by the predictions from a Gaussian Process (GP).
This provides two distinct advantages compared with contemporary approaches.
First, using a GP sidesteps the need for an accurate model-based description
of the system dynamics (or a model choice or model-averaging approach which
introduces yet more difficulty).  
Second, unlike parametric model-fitting approaches, which can only reflect 
uncertainty in the corresponding confidence intervals or  posterior distributions
of the parameters, the GP approach provides a state-space dependent representation
of uncertainty.  In other words, a GP estimated from a given data set permits 
a much greater uncertainty far from the observed data than it does near the observed 
data.

<!-- Tipping points and the frequency of a lack of data throughout state-space -->

The idea that any approach can perform well without either having to know the model or
have particularly good data should immediately draw suspicion.  The reader must bear
in mind that the strength of our approach comes not from some magic predictive power
from such limited information, but rather, by providing a more honest expression of 
uncertainty outside the observed data without sacrificing the predictive capacity 
near the observed data.

By coupling this more accurate description of what is known and unknown to the 
decision-making under uncertainty framework provided by stochastic dynamic programming,
we are able to obtain more robust management policies than with common parameteric modeling 
approaches.
